ComBatLS Replication Results

### 2024-08-21 eda/lifespan_eda.Rmd
working on retaining more output from t-tests

Load data from Jakob - updated with euler numbers!
```{r}
cn.df <- fread(paste0(base_path,"/data/lifespan_dataset_dkatlas_CNonly_euler.csv"), stringsAsFactors = TRUE)

#drop repeat scans
cn.df.unique <- cn.df %>%
  drop_na(c(age_days, participant, study)) %>%
  arrange(age_days) %>%
  group_by(participant) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(sexMale = as.numeric(case_when(sex == "Male" ~ 1,
                                        sex == "Female" ~ 0,
                                        TRUE ~ NA)),
         sex.age = sexMale*age_days,
         log_age = log(age_days, base=10))

@@

>>>>

Load data from Jakob - updated with euler numbers!
```{r}
cn.df <- fread(paste0(base_path,"/data/lifespan_dataset_dkatlas_CNonly_euler.csv"), stringsAsFactors = TRUE)
```

drop UKB for sharing
```{r}
cn.df_dropUKB <- cn.df %>%
  dplyr::filter(study != "UKB")
fwrite(cn.df_dropUKB, file=paste0(base_path, "/data/lifespan_dataset_dkatlas_CNonly_euler_noukb.csv"))
```

```{r}
#drop repeat scans
cn.df.unique <- cn.df %>%
  drop_na(c(age_days, participant, study)) %>%
  arrange(age_days) %>%
  group_by(participant) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(sexMale = as.numeric(case_when(sex == "Male" ~ 1,
                                        sex == "Female" ~ 0,
                                        TRUE ~ NA)),
         sex.age = sexMale*age_days,
         log_age = log(age_days, base=10))

@@

drop UKB for sharing
```{r}
cn.df.imp_dropUKB <- cn.df.imp %>%
  dplyr::filter(study != "UKB")
fwrite(cn.df.imp_dropUKB, file=paste0(base_path, "/data/lifespan_CN_imp-sites_euler_noukb.csv"))
```

### 2024-08-21 paper_replication_results.Rmd
working on retaining more output from t-tests

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, gdata, gtExtras, webshot2)
#Sys.setenv("MC_CORES"=10L)
options("mc.cores")

source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
source("./cubic_scripts/R_scripts/ranked_welch_tests.R")
aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
```

@@

t.test.cent.sex.test.med <- cent.sex.test.med %>%
  group_by(dataset, pheno_cat) %>%
  summarise(t.test.p = tidy(t.test(median_of_median_sex_diff))$p.value,
            t.test.est = tidy(t.test(median_of_median_sex_diff))$estimate,
            x_pos = median(median_of_median_sex_diff)) %>% #test median of each distribution against 0
  mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = length(unique(cent.sex.test.med$dataset))*length(unique(cent.sex.test.med$pheno_cat))),
                  sig_fdr = case_when(p.val_fdr < 0.001 ~ "***",
                               p.val_fdr >= 0.001 & p.val_fdr < 0.01 ~ "**",
                               p.val_fdr >= 0.01 & p.val_fdr < 0.05 ~ "*",
                               p.val_fdr >= 0.05 ~ ""),
         y_pos = case_when(dataset=="cf" ~ 9.3,
                           dataset=="cf.lm" ~ 11,
                           dataset=="cf.gam" ~ 10.5,
                           dataset=="cf.gamlss" ~ 6))
#get vals
t.test.cent.sex.test.med

@@

t.test.z.sex.test.med <- z.sex.test.med %>%
  group_by(dataset, pheno_cat) %>%
  summarise(t.test.p = tidy(t.test(median_of_median_sex_diff))$p.value,
            t.test.est = tidy(t.test(median_of_median_sex_diff))$estimate,
            x_pos = median(median_of_median_sex_diff)) %>% #test median of each distribution against 0
  mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = length(unique(z.sex.test.med$dataset))*length(unique(z.sex.test.med$pheno_cat))),
                  sig_fdr = case_when(p.val_fdr < 0.001 ~ "***",
                               p.val_fdr >= 0.001 & p.val_fdr < 0.01 ~ "**",
                               p.val_fdr >= 0.01 & p.val_fdr < 0.05 ~ "*",
                               p.val_fdr >= 0.05 ~ ""),
         y_pos = case_when(dataset=="cf" ~ 155,
                           dataset=="cf.lm" ~ 360,
                           dataset=="cf.gam" ~ 345,
                           dataset=="cf.gamlss" ~ 125))

t.test.z.sex.test.med.filt <- t.test.z.sex.test.med %>% 
  dplyr::filter(dataset != "cf")

@@

```{r}
fig_3_wtest <- ggarrange(
  ggarrange(
  ggarrange(sex.var.plot.global, sex.var.plot,
  ncol=2, nrow=1, 
  widths = c(1, 3.25),
  hjust=0.01,
  align = "hv",
  common.legend = TRUE, legend = "bottom"
  ),
  sex_bias_density_plt,
  nrow=2, ncol=1,
  labels=c("A", "B")
  ),
  all_cent_sum_boxplt_wtest,
  ncol=2, nrow=1,
  labels=c(NA, "C"),
  widths = c(1, 1.8)
  # #align = "hv",
  # heights=c(1, 1.8),
  # common.legend = FALSE,
  # #labels=c("A", "B")
  )

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, ggseg, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, gdata, gtExtras, webshot2) #ggsegExtra
#Sys.setenv("MC_CORES"=10L)
options("mc.cores")

source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
source("./cubic_scripts/R_scripts/ranked_welch_tests.R")
remotes::install_github("BGDlab/ggseg-tissueclass-atlas")
```

@@

#t-test of median sex diffs in centile error distributions vs 0
t.test.cent.sex.test.med <- cent.sex.test.med %>%
  group_by(dataset, pheno_cat) %>%
  summarise(t.test.p = tidy(t.test(median_of_median_sex_diff))$p.value,
            t.test.est = tidy(t.test(median_of_median_sex_diff))$estimate,
            t.test.stat = tidy(t.test(median_of_median_sex_diff))$statistic,
            t.test.df = tidy(t.test(median_of_median_sex_diff))$parameter,
            x_pos = median(median_of_median_sex_diff)) %>% #test median of each distribution against 0
  mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = length(unique(cent.sex.test.med$dataset))*length(unique(cent.sex.test.med$pheno_cat))),
                  sig_fdr = case_when(p.val_fdr < 0.001 ~ "***",
                               p.val_fdr >= 0.001 & p.val_fdr < 0.01 ~ "**",
                               p.val_fdr >= 0.01 & p.val_fdr < 0.05 ~ "*",
                               p.val_fdr >= 0.05 ~ ""),
         y_pos = case_when(dataset=="cf" ~ 9.3,
                           dataset=="cf.lm" ~ 11,
                           dataset=="cf.gam" ~ 10.5,
                           dataset=="cf.gamlss" ~ 6))
#get vals
t.test.cent.sex.test.med_tbl <- t.test.cent.sex.test.med %>%
  dplyr::select(dataset, pheno_cat, t.test.est, t.test.stat, t.test.df, p.val_fdr, sig_fdr) %>%
  mutate(dataset = case_when(dataset == "cf" ~ "ComBat w/o Covariates",
                                   dataset == "cf.lm" ~ "Linear ComBat",
                                   dataset == "cf.gam" ~ "ComBat-GAM",
                                   dataset == "cf.gamlss" ~ "ComBatLS")) %>%
  rename("ComBat Method" = dataset,
         "Phenotype Category" = pheno_cat,
         "Estimate" = t.test.est,
         "t-statistic" = t.test.stat,
         "dfs" = t.test.df,
         "p-value (FDR-corrected)" = p.val_fdr,
         "Significance" = sig_fdr) %>%
  gt() %>%
  tab_style(style = list(cell_fill(color = "gray95")),
    locations = cells_row_groups()) %>%
  tab_options(column_labels.font.weight = "bold") %>%
  tab_header(title = "One-sided t-tests of sex differences in centile errors across 100 replications")
gtsave(t.test.cent.sex.test.med_tbl, filename="figs/cent_median_sexdiff_table.png")
```

plots
```{r}
t.test.cent.sex.test.med %>%
  ggplot() +
  geom_col(aes(x=pheno_cat, y=sig_fdr, fill=dataset), position="dodge")

@@

t.test.z.sex.test.med <- z.sex.test.med %>%
  group_by(dataset, pheno_cat) %>%
  summarise(t.test.p = tidy(t.test(median_of_median_sex_diff))$p.value,
            t.test.est = tidy(t.test(median_of_median_sex_diff))$estimate,
            t.test.stat = tidy(t.test(median_of_median_sex_diff))$statistic,
            t.test.df = tidy(t.test(median_of_median_sex_diff))$parameter,
            x_pos = median(median_of_median_sex_diff)) %>% #test median of each distribution against 0
  mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = length(unique(z.sex.test.med$dataset))*length(unique(z.sex.test.med$pheno_cat))),
                  sig_fdr = case_when(p.val_fdr < 0.001 ~ "***",
                               p.val_fdr >= 0.001 & p.val_fdr < 0.01 ~ "**",
                               p.val_fdr >= 0.01 & p.val_fdr < 0.05 ~ "*",
                               p.val_fdr >= 0.05 ~ ""),
         y_pos = case_when(dataset=="cf" ~ 155,
                           dataset=="cf.lm" ~ 360,
                           dataset=="cf.gam" ~ 345,
                           dataset=="cf.gamlss" ~ 125))

#pretty table
t.test.z.sex.test.med_tbl <- t.test.z.sex.test.med %>%
  dplyr::select(dataset, pheno_cat, t.test.est, t.test.stat, t.test.df, p.val_fdr, sig_fdr) %>%
  mutate(dataset = case_when(dataset == "cf" ~ "ComBat w/o Covariates",
                                   dataset == "cf.lm" ~ "Linear ComBat",
                                   dataset == "cf.gam" ~ "ComBat-GAM",
                                   dataset == "cf.gamlss" ~ "ComBatLS")) %>%
  rename("ComBat Method" = dataset,
         "Phenotype Category" = pheno_cat,
         "Estimate" = t.test.est,
         "t-statistic" = t.test.stat,
         "dfs" = t.test.df,
         "p-value (FDR-corrected)" = p.val_fdr,
         "Significance" = sig_fdr) %>%
  gt() %>%
  tab_style(style = list(cell_fill(color = "gray95")),
    locations = cells_row_groups()) %>%
  tab_options(column_labels.font.weight = "bold") %>%
  tab_header(title = "One-sided t-tests of sex differences in z-score errors across 100 replications")
gtsave(t.test.z.sex.test.med_tbl, filename="figs/z_median_sexdiff_table.png")
```

plot
```{r}
t.test.z.sex.test.med.filt <- t.test.z.sex.test.med %>% 
  dplyr::filter(dataset != "cf")

@@

```{r}
fig_3_wtest <- ggarrange(
  ggarrange(
  ggarrange(sex.var.plot.global, sex.var.plot,
  ncol=2, nrow=1, 
  widths = c(1, 3.25),
  hjust=0.01,
  align = "hv",
  common.legend = TRUE, legend = "bottom"
  ),
  sex_bias_density_plt,
  nrow=2, ncol=1,
  labels=c("A", "B")
  ),
  all_cent_sum_boxplt.diff,
  ncol=2, nrow=1,
  labels=c(NA, "C"),
  widths = c(1, 1.8)
  # #align = "hv",
  # heights=c(1, 1.8),
  # common.legend = FALSE,
  # #labels=c("A", "B")
  )
